---
title: "Sharing rules across packages"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Sharing rules across packages}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(flir)
```

Besides helping to detect a built-in list of bad patterns in R code, `flir` can also be used to refactor code, for instance by replacing the use of [deprecated or superseded](https://lifecycle.r-lib.org/articles/stages.html) functions.
If you work on a project that uses `dplyr`, you could for instance [write your own rules](https://flir.etiennebacher.com/articles/adding_rules) to replace the use of `sample_n()` by `slice_sample()` (as described in [this blog post](https://www.etiennebacher.com/posts/2025-05-23-refactoring-code-with-flir/)).

However, that puts the burden of writing such rules on the end user, who is not necessarily familiar with all the changes in the package they use. 
Therefore, `flir` provides a way for package developers to define their own rules in the package so that anyone who uses it will also have access to the same set of rules.

Let's take an example with one of my packages: [`tidypolars`](https://tidypolars.etiennebacher.com/).

## State of the package

At the time of writing, `tidypolars` (0.13.0) contains several deprecated functions, such as `describe()` and `describe_plan()`. Those functions should be replaced by `summary()` and `explain(optimized = FALSE)` respectively.

## For package developers

On the side of package developers (in this example, this is only me), we can define some rules for `tidypolars` and store them in `inst/flir/rules`^[I won't spend time explaining each part of the file, refer to the [Adding rules](https://flir.etiennebacher.com/articles/adding_rules) vignette for more details.].

First, I define a rule to detect the use of `describe()` and replace it with `summary()`:

```yaml
id: deprecated-describe
language: r
severity: warning
rule:
  pattern: describe($$$ARGS)
fix: summary(~~ARGS~~)
message: In tidypolars, `describe()` is deprecated. Use `summary()` instead.
```

Keep in mind that `flir` works mainly by using *code patterns*, it doesn't have information on the R environment.
Therefore, the rule above doesn't know from which package the function comes from and it would also replace `describe()` calls coming from other packages.

I do the same with `describe_plan()`:

```yaml
id: deprecated-describe_plan
language: r
severity: warning
rule:
  pattern: describe_plan($ARG)
fix: explain(~~ARG~~, optimized = FALSE)
message: In tidypolars, `describe_plan()` is deprecated. Use `explain(optimized = FALSE)` instead.
```

This is all that is needed from package developers. When users will install the package, they will have access to those rules.

**Note:** it is good practice to have a single name for a family of rules and to use this name as filename. 
For instance, the first rule has `id: deprecated-describe`, so I should name the file `deprecated-describe.yml`.
If you want to store several rules related to this deprecation, you could name the other rules `deprecated-describe_2`, `deprecated-describe_3`, etc. and store them in the same file, separated by `...`.

## For users

To use rules defined in other packages, you need to set up a `flir` folder with `setup_flir()`. 
This will create the file `flir/config.yml` that we will modify to get rules from other packages.

Once this is done, you can add a field `from-package` and list the packages from which you want to import the rules. For instance, to benefit from the rules defined above, we can do:

```yaml
from-package:
  - tidypolars
```

and that's it! From now on, all calls to `lint_*()` or `fix_*()` on the project will also use the rules stored in `tidypolars`.
Note that those rules are bundled with the package, meaning that if the package developers change or remove some rules and you update the package, the rules you will have access to will be different.

As before, you can also list rules to exclude in the `exclude` field of `flir/config.yml` but you will need to prefix them with `from-<pkgname>-`. 
For instance, to ignore the rule about the deprecation of `describe()`, you can set:

```yaml
exclude:
  - from-tidypolars-deprecated-describe
```
